---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Loading libraries:

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
```

The following settings are local-environment dependent so they have to be 
adjusted according to the local computer settings.

```{r, echo=TRUE}

# Folder with the cloned repository
project_folder = "D:/projects/RepData_PeerAssessment1"

```

Code responsible for unzipping the archive and reading the data:

```{r, echo=TRUE}

original_folder = getwd()
setwd(project_folder)

unzip("activity.zip", overwrite = TRUE, exdir = ".")

original_data <- read_delim("activity.csv", 
                           skip = 1,
                           delim = ",",
                           col_names = c("steps", "date", "interval"),
                            col_types = cols(
                              steps = col_integer(),
                              date = col_date(format= "%Y-%m-%d"),
                              interval = col_factor()))

summary(original_data)
```


## What is mean total number of steps taken per day?

The histogram for the "total number of steps per day" distribution

```{r sum-steps-histogram, echo = TRUE}

total_steps_per_day <- original_data %>% select(steps, date) %>% 
                      group_by(date) %>%
                      summarise(sum_steps_per_day = sum(steps, na.rm = TRUE)) 
                      
ggplot(data = total_steps_per_day, aes(x = sum_steps_per_day)) + 
      geom_histogram(bins = 30)

```



Calculating the mean and the median of the total number of steps taken per day:

```{r, echo=TRUE}
mean_median_total_steps <- total_steps_per_day %>%
  summarize(mean = mean(sum_steps_per_day), median = median(sum_steps_per_day))
mean_median_total_steps
```

The mean of the total number of steps per day is `r mean_median_total_steps$mean`
and the median -  `r mean_median_total_steps$median`.

## What is the average daily activity pattern?

```{r average_daily_activities, echo=TRUE}

mean_day_activities <- original_data %>% group_by(interval) %>%
                       summarize(average_activity = mean(steps, na.rm = TRUE)) %>%
                       mutate(interval = as.numeric(as.character(interval))) %>%
                       arrange(interval)

ggplot(mean_day_activities, aes(x=interval, y=average_activity)) +
geom_line()
 
```


The interval `r mean_day_activities[which(mean_day_activities$average_activity == max(mean_day_activities$average_activity) ), ]$interval`  contains on the 
average the biggest number of steps done throughout the day.


## Imputing missing values

```{r rows_with_NAs, echo=TRUE}
number_rows_with_na <- sum(is.na(original_data$steps) | is.na(original_data$date) | is.na(original_data$interval))
```

There are `r number_rows_with_na` rows with missing data in the original dataset.
The propose imputing strategy is to use the mean number of steps form the specified interval from all days. 

The imputed data is stored in the original_imputed_data. 
```{r echo=TRUE}
original_imputed_data <- original_data %>%
                         mutate(interval = as.integer(as.character(interval))) %>%
                         mutate(steps = as.double(steps))
original_imputed_data <- left_join(original_imputed_data, mean_day_activities, by="interval")

original_imputed_data <- original_imputed_data %>% 
                         mutate(steps = ifelse(is.na(steps), average_activity, 
                                               steps)) %>%
                         select(-average_activity)
                         
```

The new histogram of the total steps per day is as it follows:

```{r  sum-steps-histogram-imputed, echo = TRUE}


total_imputed_steps_per_day <- original_imputed_data %>% select(steps, date) %>% 
                      group_by(date) %>%
                      summarise(sum_steps_per_day = sum(steps, na.rm = TRUE)) 
                      
ggplot(data = total_imputed_steps_per_day, aes(x = sum_steps_per_day)) + 
      geom_histogram(bins = 30)
```

Computing the mean / median of total steps per day: 

```{r, echo=TRUE}
mean_median_total_steps_imputed <- total_imputed_steps_per_day %>%
  summarize(mean = mean(sum_steps_per_day), median = median(sum_steps_per_day))
mean_median_total_steps_imputed
```


The mean of the total number of steps per day is `r format(mean_median_total_steps_imputed$mean,2)`
and the median -  `r format(mean_median_total_steps_imputed$median,2)` (all values are rounded). After data imputing
those values are naturally higher. 

## Are there differences in activity patterns between weekdays and weekends?
