---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Loading libraries:

```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
```

The following settings are local-environment dependent so they have to be 
adjusted according to the local computer settings.

```{r, echo=TRUE}

# Folder with the cloned repository
project_folder = "D:/projects/RepData_PeerAssessment1"

```

Code responsible for unzipping the archive and reading the data:

```{r, echo=TRUE}

original_folder = getwd()
setwd(project_folder)

unzip("activity.zip", overwrite = TRUE, exdir = ".")

original_data <- read_delim("activity.csv", 
                           skip = 1,
                           delim = ",",
                           col_names = c("steps", "date", "interval"),
                            col_types = cols(
                              steps = col_integer(),
                              date = col_date(format= "%Y-%m-%d"),
                              interval = col_factor()))

summary(original_data)
```


## What is mean total number of steps taken per day?

The histogram for the "total number of steps per day" distribution

```{r sum-steps-histogram, echo = TRUE}

total_steps_per_day <- original_data %>% select(steps, date) %>% 
                      group_by(date) %>%
                      summarise(sum_steps_per_day = sum(steps, na.rm = TRUE)) 
                      
ggplot(data = total_steps_per_day, aes(x = sum_steps_per_day)) + 
      geom_histogram(bins = 30)

```



Calculating the mean and the median of the total number of steps taken per day:

```{r, echo=TRUE}
mean_median_total_steps <- total_steps_per_day %>%
  summarize(mean = mean(sum_steps_per_day), median = median(sum_steps_per_day))
mean_median_total_steps
```

The mean of the total number of steps per day is `r mean_median_total_steps$mean`
and the median -  `r mean_median_total_steps$median`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
